### 2017-07-21 sheets/talk-11/sheet-11.tex

\begin{exercise}
  Give the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
  the presheaf model.
\end{exercise}
\begin{answer}
  \[
    \mathit{Ty}_F(\Gamma) := \mathit{Psh(\Gamma)}
  \]
  \[
    \mathit{Ter}_F(\Gamma, A) := \{(a\rho \in A\rho)_{I \in \cat{C}, \rho \in \Gamma(I)} | \forall f \from J \to I, a\rho f = a(\rho f) \}
  \]
\end{answer}

@@

  We can decompose $\sigma'$ as $\sigma' = (p_\Gamma, q_\Gamma) \circ \sigma' =
  (p_\Gamma \circ \sigma', q_\Gamma\sigma') = (\sigma \circ p', M)$, where $M
  \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash A(\sigma \circ p')
  ~\text{type}$). Let $\delta$ be the morphism
  \[
    \delta \equiv (p', M) \from \Delta' \to \Delta.A\sigma
  \]
  which is suitable since $p_\Delta \circ \delta = p'$ and
  \[
    (\sigma \circ p_\Delta, q) \circ \delta =
    (\sigma \circ p_\Delta \circ \delta, q\delta) =
    (\sigma \circ p', q_\Gamma\sigma') =
    (p_\Gamma \circ \sigma', q_\Gamma\sigma') =
    (p_\Gamma, q_\Gamma) \circ \sigma' =
    \sigma'
  \]

@@

  Let's now see that $\int_\cat{C}\sigma$ is a functor in $\CAT$. We have a
  mapping for objects $\int_\cat{C}\sigma(I, \rho)$ given above. The operation
  on arrows is as follows, where $f \from J \to I$, $g \from K \to J$:
  \[
    \int_\cat{C}\sigma(f) := \sigma_{I \to J} \Gamma_f
  \]
  Then, keeping in mind that composition $fg$ is applied on the right we get
  $\int_\cat{C}\sigma(fg) = \sigma_{I \to K} \Gamma_{fg} = \sigma_{I \to K}
  \lambda \rho.\rho (fg) = \sigma_{I \to K} \lambda \rho. (\rho f)g = \sigma_{J
  \to K} (\sigma_{I \to J} \lambda \rho.\rho f) g = (\sigma_{J \to K} \lambda
  \rho.\rho g) (\sigma_{I \to J} \lambda.\rho f) = (\sigma_{J \to K} \Gamma_g)
  (\sigma_{I \to J} \Gamma_f)$. Which, together with $\rho 1 = \rho$ confirms
  that $\int_\cat{C}\sigma$ is indeed a functor.

@@

  \emph{Hint}: Read section $1.2.1$ in Huber's thesis for the general case.
\end{exercise}
\begin{answer}
  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we give the
  element of $\mathit{Ty}(\Gamma)$ corresponding to the type $A \to A$:
  \[
    (A \to A)\rho = \{w | w_f \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
  \]
  The required equalities hold trivially for any $g \from K \to J$, $((\lambda
  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.

@@

  Now suppose that we have dependent types $\Gamma \vdash C \text{ type}$ and
  $\Gamma, C \vdash D \text{ type}$.

@@

  \emph{Hint}: The definition is \emph{almost} identical to the one given
  above. But what do you need to change? To recall the presheaf model, read
  section $1.2$ in Huber's thesis.
\end{exercise}
\begin{answer}
  \[
    (\Sigma AB)\rho := \{(a,b) | a \in A\rho \land b \in B(\rho, a)\}
  \]
\end{answer}

@@

  Now to get an inhabitant of that type let's use $\mathit{Subst}$:
  \[
    P : Id(0, Suc(0)) \vdash \mathit{Subst}(P, El(f(0))) : El(f(Suc(0)))
  \]
  Since the inhabitant to $El(f(0)) = El(\hat{1}) = 1~\text{type}$, namely
  $\star$, is always present by the formation rules, and $El(f(Suc(0))) =
  0~\text{type}$ we obtain an inhabitant to the required type.
\end{answer}

>>>>

\begin{exercise}
  Give the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
  the presheaf model.
\end{exercise}
\begin{answer}
  \[
    \mathit{Ty}_F(\Gamma) := \mathit{Psh(\int_\cat{C} \Gamma)}
  \]
  \[
    \mathit{Ter}_F(\Gamma, A) := \{(a\rho \in A\rho)_{I \in \cat{C}, \rho \in \Gamma(I)} | \forall f \from J \to I, a\rho f = a(\rho f) \}
  \]
\end{answer}

@@

  We can decompose $\sigma'$ as $\sigma' = 1 \circ \sigma' = (p_\Gamma,
  q_\Gamma) \circ \sigma' = (p_\Gamma \circ \sigma', q_\Gamma\sigma') = (\sigma
  \circ p', M)$, where $M \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash
  A(\sigma \circ p') ~\text{type}$). Let $\delta$ be the morphism
  \[
    \delta \equiv (p', M) \from \Delta' \to \Delta.A\sigma
  \]
  which is suitable since $p_\Delta \circ \delta = p'$ and
  \[
    (\sigma \circ p_\Delta, q) \circ \delta =
    (\sigma \circ p_\Delta \circ \delta, q\delta) =
    (\sigma \circ p', q_\Gamma\sigma') =
    (p_\Gamma \circ \sigma', q_\Gamma\sigma') =
    (p_\Gamma, q_\Gamma) \circ \sigma' =
    1 \circ \sigma' =
    \sigma'
  \]

@@

  Let's now see that $\int_\cat{C}\sigma$ is a functor in $\CAT$. We have a
  mapping for objects $\int_\cat{C}\sigma(I, \rho)$ given above. The operation
  on arrows is as follows, where $f \from J \to I$, $g \from K \to J$:
  \[
    \int_\cat{C}\sigma(f) := \sigma_{I \to J} \Gamma_f
  \]
  Recall the naturality axiom for the natural transformation $\sigma$. It
  requires for each $I \in \cat{C}$ a corresponding $\sigma_I$. Hence the
  operation of $\sigma$ on the composition of morphisms in $\cat{C}$ can be
  considered as application of two indexed $\sigma_i$ of the corresponding
  objects between the morphisms (refer to the naturality square to see this).
  Below we give the indices on $\sigma$ that correspond to the place in the
  naturality square for clarity.  Then with associativity and keeping in mind
  that the composition $fg$ and the individual $f$ and $g$ are applied on the
  right we get $\int_\cat{C}\sigma(fg) = \sigma_{I \to K} \Gamma_{fg} =
  \sigma_{I \to K} \lambda \rho.\rho (fg) = \sigma_{I \to K} \lambda \rho.
  (\rho f)g = \sigma_{J \to K} (\sigma_{I \to J} \lambda \rho.\rho f) g =
  (\sigma_{J \to K} \lambda \rho.\rho g) (\sigma_{I \to J} \lambda \rho.\rho f)
  = (\sigma_{J \to K} \Gamma_g) (\sigma_{I \to J} \Gamma_f)$.  Which, together
  with $\rho 1 = \rho$ confirms that $\int_\cat{C}\sigma$ is indeed a functor.

@@

  \emph{Hint}: Read section $1.2.1$ in Huber's thesis for the general case.
\end{exercise}
\begin{answer}
  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we give the
  element of $\mathit{Ty}(\Gamma)$ corresponding to the type $A \to A$:
  \[
    (A \to A)\rho = \{w | w_f u \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
  \]
  The required equalities hold trivially for any $g \from K \to J$, $((\lambda
  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.

@@

  Now suppose that we have dependent types $\Gamma \vdash A \text{ type}$ and
  $\Gamma, A \vdash B \text{ type}$.

@@

  \emph{Hint}: The definition is \emph{almost} identical to the one given
  above. But what do you need to change? To recall the presheaf model, read
  section $1.2$ in Huber's thesis.
\end{exercise}
\begin{answer}
  \[
    (\Sigma AB)\rho := \{(a,b) | a \in A\rho \land b \in B(\rho, a)\}
  \]
  where $(a, b)f = (af, bf)$ for $f \from J \to I$. The pairing operation
  $\Gamma \vdash (u, v) : \Sigma AB$ of $\Gamma \vdash u : A$ and $\Gamma.A
  \vdash v : B(1, u)$ is defined by the componentwise pairing: $(u, v)\rho =
  (u\rho, v\rho)$. Likewise for the projections $p$ and $q$, $(pw)\rho = a$ and
  $(qw)\rho = b$. This validates the necessary equations.
\end{answer}

@@

  Now to get an inhabitant of that type let's use $\mathit{Subst}$:
  \[
    P : Id(0, Suc(0)) \vdash \mathit{Subst}(P, (\star : El(f(0)))) : El(f(Suc(0)))
  \]
  Since the inhabitant to $El(f(0)) = El(\hat{1}) = 1~\text{type}$, namely
  $\star$, is always present by the formation rules, and $El(f(Suc(0))) =
  0~\text{type}$ we obtain an inhabitant to the required type.
\end{answer}
